\documentclass{article}
\usepackage{textcomp, gensymb}
\usepackage{utf8add}
\usepackage[most]{tcolorbox}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tcolorbox}
\usepackage{outlines}
\usepackage{enumitem}
\usepackage{graphics}
\usepackage{bbm}
% \usepackage{hhline}

\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% \setenumerate[1]{label=\Roman*.}
% \setenumerate[2]{label=\Alph*.}
% \setenumerate[3]{label=\roman*.}
% \setenumerate[4]{label=\alph*.}

\title{CMPUT 428: 3D Modelling}
\author{Roderick Lan}
% \date{January 11, 2024}
\date{}

\usepackage{natbib}
\makeatletter
% \crefformat{tcb@cnt@Example}{example~#2#1#3}
% \Crefformat{tcb@cnt@Example}{Example~#2#1#3}
\makeatother
\newtcbtheorem[auto counter, number within = subsection]
{definition}{Definition}{%                                                        
  breakable,
  fonttitle = \bfseries,
  colframe = blue!75!black,
  colback = blue!10
}{def}

\makeatother
\newtcbtheorem[auto counter, number within = subsection]
{example}{Example}{%                                                        
  breakable,
  fonttitle = \bfseries,
  colframe = orange!75!black,
  colback = orange!10
}{ex}

\makeatother
\newtcbtheorem[auto counter, number within = subsection]
{expln}{Expln}{%                                                        
  breakable,
  fonttitle = \bfseries,
  colframe = red!75!black,
  colback = red!10
}{exp}

\makeatother
\newtcbtheorem[auto counter, number within = subsection]
{ovr}{Overview}{%                                                        
  breakable,
  fonttitle = \bfseries,
  colframe = blue!75!black,
  colback = blue!10!white!50
}{ovr}


\makeatother
\newtcbtheorem[auto counter, number within = subsection]
{refer}{Reference}{%                                                        
  breakable,
  fonttitle = \bfseries,
  colframe = red!75!black,
  colback = red!10
}{refer}
\begin{document}

\makeatother
\newtcbtheorem[auto counter, number within = subsection]
{thm}{Thm}{%                                                        
  breakable,
  fonttitle = \bfseries,
  colframe = orange!75!black,
  colback = orange!10
}{thm}

% \setcounter{section}{1}
\maketitle
\tableofcontents
\break
% \section*{Lecture 1}

\section{Lecture - Feb 29}
lec07modeling18EmbMedia.pdf\\[5pt]
3D from video is:
\begin{outline}
    \1 inexpensive
    \1 quick and convenient for the user
    \1 integrates with existing software (ie. maya, blender, c4d, ue4, etc.)
\end{outline}
\subsection{Prelims; Macro Geometry}
Shape from silhouettes:
\begin{list}{}{}
    \item more = better
    \item provides good starting approx. for object
\end{list}
Structure from motion: % photogrammetry
\begin{list}{}{}
    \item uses triangulations to get 3d points
    \item can be done with 'unstructured' imgs/scene (ie. from phone camera)
    % \item 
\end{list}


\subsection{Fundamental Problems}
\begin{enumerate}
    \item \textbf{Resection} of 3D camera matrices
    \item \textbf{Intersection}
    \item \textbf{Structure From Motion (SFM)}
\end{enumerate}

\subsubsection{Resection}
Projection Equation:
\[
    x_i = P_i X
\]
Resection:
\[
    x_i, X\to P
\]
Imagine you have $x$ (ie. 3d model) and image $x_i$ (projection of points on img). 
Finding camera = finding projection of rays that map 3d point to image point.
\\
Can have many potential camera planes that all map (?)


\subsubsection{Intersection}
Projection Equation:
\[
    x_i = P_i X
\]
Intersection:
\[
    x_i, P_i \to X
\]
Have $\ge 2$ cameras; projections in $\ge 2$ images. Intersect rays to get 3d points. 
\\
Need to 'know' the camera properties. 

\subsubsection{SFM}
Projection Equation:
\[
    x_i = P_i X
\]
Structure From Motion (SFM):
\[
    x_i \to P_i, X
\]
Get image predictions ($x_i$) from point tracking.
\\
Get image points in $\ge 2$ views, calculate 3d points (structure) and 
camera projection matrices (motion)
\\
Estimate projective structure\\
Rectify the reconstruction to metric (autocalibration)
\\
(dont know the cameras; figure out constraints)

\subsection{Seq. 3D Structure From Motion}
Use 2 and 3 view geometry
\begin{outline}[enumerate]
    \1 Initialize structure and motion from 2 views
    \1 for each additional view:
        \2 determine pose
        \2 refine and extend structure
    \1 determine correspondences robustly by jointly estimating matchies and epipolar constraints
\end{outline}
\subsection{2 View Geometry}
2 points in views $x_i$, comes from 3d point $X$. If you have camera center, know that ray 
from center to point extends to infinity; $X$ lies somewhere on ray. 
\\
line connecting camera centers
\\
epipole - intersecting point b/w line and image plane
\\
epipolar line - line b/w imaged $x_i$ and epipole
\subsubsection{Epipolar Geometry}
C, C' x, x' and X are coplanar
\\
Epipolar geometry exists independent of 3d point, only related to camera and 2d point
projections.
\\
% Dont know $X$, argue that 
There is a family of epipolar planes \\
Epipolar lines go towards parallel lines
\\
Family of planes and lines intersect infinitely far apart.
\\
Epipoles ($e, e'$) can be finite or infinite
\begin{list}{}{}
    \item intersection of baseline with img plane
    \item projection of projection center in other img
    \item vanishing point of camera motion direction
\end{list}
Epipolar plane - plane containing baseline (1D family)
Epipolar line - intersection of epipolar plane with image 

\begin{example}
    {Parallel Motion with img plane}{}
    Epipoles at infinity
\end{example}
% optic flow must lie on epipolar lines
% \begin{example}
%     {Forward motion}{}

% \end{example}

\subsection{Fundamental Matrix}
Algebraic representation of epipolar geometry.
\[
    \mathbf x \mapsto \mathbf l'
\]
This mapping is a singular correlation; represented by fundamental matrix $F$
\\
$3 \times 3$ matrix with 7 dof (8 independent ratios)

\subsubsection{Algebraic Derivation}
Let $x$ be $PX$
\begin{flalign*}
    P^{+}x &= P^{+}P X
    \intertext{Say that pseudoinverse represents line from center to point}
    \mathbf X(\lambda) &= \mathbf P^+ \mathbf x + \lambda \mathbf C \\
    \mathbf l' &= \mathbf{P'C} \times \mathbf{P'P^+x} \\
    F &= [e']_\times P'P^+
\end{flalign*}
$\lambda$ is a scalar
\\
$[e']_\times$ skew symmetric cross product matrix (rank 2)
\\
doesnt work for $C=C' \Rightarrow F=0$ (must have distinct camera centers)
\\[5pt]
Alternatively can write:
\[
    F = [e']_\times H_\infty
\]
\[
    H_\infty = K^{-1}RK
\]
\subsubsection{Geometric Derivation}
Step 1: $X$ on a plane $\pi$
\[
    x' = H_\pi x
\]
Step 2: Epipolar line $l'$ 
\[
    l' = e' \times x' = [e']_\times H_\pi x = Fx
\]
Map from 2D to 1D family (rank 2)
\\
(plane $\pi$ can be different for each point, \textbf{unrelated} to epipolar plane)

\subsubsection{Correspondence Condition}
Exists 2D matrix that ouputs epipolar line.
\\
If $\mathbf x'$ lies on $\mathbf l'$, $\mathbf{x' l'} = 0$ 
\[
    \mathbf{x' \cdot Fx} = 0
\]

\noindent\rule{\textwidth}{.5pt}
\\[15pt]
$\mathbf F$ is the unique $3 \times 3$ rank 2 matrix that satisfies 
\[
    \mathbf{x'^\top Fx} = 0 \ \ \ \ \forall \mathbf x\leftrightarrow \mathbf x'
\]
\begin{enumerate}
    \item \textbf{Transpose}: if $\mathbf F$ is a fundamental matrix for (P, P'), $\mathbf F^\top$
        is fundamental matrix for (P', P)
    \item Epipolar Lines: $\mathbf{l' = Fx}$ and $\mathbf{l=F^\top x'}$
    \item Epipoles: on all epipolar lines, $\mathbf{e'^\top Fx = 0}$ $\forall \mathbf x, 
    \Rightarrow e'^\top F=0$, similarly $\mathbf{Fe}= 0$
    \item $\mathbf F$ has 7dof ($3\times 3$ - 1 (homog), - 1 (rank 2))
    \item $\mathbf F$ is a correlation, projective mapping from a point $\mathbf x$ 
        to a line $\mathbf{l' = Fx}$ (not a proper corr., ie. not invertible)
\end{enumerate}

\subsubsection{Summary}
Algebraic rep. of epipolar geom.
\\
Step 1: $\mathbf X$ on plane $\pi$
\[
    \mathbf{x' = Hx}
\]
Step 2: epipolar line $\mathbf l'$
\begin{flalign*}
    \mathbf{l'} &= \mathbf{e' \times x' = [e']_\times x'} \\ &= \mathbf{[e']_\times Hx = Fx}
\end{flalign*}
\[
    \mathbf{x'^\top Fx =0 }
\]

\subsubsection{Computing F}
Use 8 points
\[
    \mathbf{x'^\top Fx} =0 
\]
(SLIDE 43-44)



\section{Lecture - Mar 5}
\subsection{Fundamental Matrix (Summary)}
get ray from point, real point could be any point that lies on ray, take 
multiple imgs and get intersection of rays to find point.
\\
% calibrated image - know location of camera
find mathematical nature of point b/w 2 views (?)

\begin{definition}
    {$F$}{}
    $3 \times 3$ Rank 2, $\mathrm{det}(F)=0$\\
    Linear solution - 8 corr. points (unique)
    \\
    Nonlinear sol - 7 corr. points (3 sol)
    \\
    Very sensitive to \textbf{noise} and \textbf{outliers} (need to know $x$ and $x'$ very precisely)
\end{definition}
\noindent
Epipolar Lines:
\[
    \mathbf l' = F\mathbf x, \ \mathbf l = F^\top \mathbf x'
\]
Epipoles:
\[
    F\mathbf e = 0, F^\top \mathbf e' = 0
\]
Projection Matrices:
\[
    P = \begin{bmatrix}
        I & | & \mathbf 0
    \end{bmatrix}
\]
\[
    P' = \begin{bmatrix}
        [\mathbf e']_\times F + \mathbf{e'v^\top} & | & \lambda \mathbf e'
    \end{bmatrix}
\]



\subsection{3D geom and 2D images w/ F}
$F$ relates to 3 questions:
\begin{enumerate}
    \item \textbf{Correspondence geometry} - given img point $x$ in first view, how does 
    this constrain the position of the corresponding point $x'$ in the second image?
    \item \textbf{Camera geometry (motion)} - given a set of corresponding img points $\{ x_i \leftrightarrow x_i' \}, i=1,\ldots,n$
    what are the cameras $P$ and $P'$ for the two views
    \item \textbf{Scene geometry (structure)} - given corresponding image points $x_i \leftrightarrow x_i'$ and 
    cameras $P, P'$, what is the position of (their pre-image) $X$ in space
\end{enumerate}
\subsection{Compute F}
(slide 43-44)
\\
should normalize pixel values [$0,\sqrt2$]
\\
Trifocal tensor - 3 view geom\\
Quadrifocal tensor - 4 view geom\\
Multiview tensors\\
(no additional constraint b/w $>4$ imgs, all constraints can be expressed
using $F$, triliear tensor or quadrifocal tensor)


\subsection{Canonical Cameras given F}
$F$ matrix corresponds to $P, P'$ iff $P'^\top FP$ is skew symmetric
\[
    X^\top P'^\top FPX = 0, \forall X
\]
$F$ matrix, $S$ skew symmetric matrix
\[
    P = \begin{bmatrix}
        I & | & 0
    \end{bmatrix}, P' = \begin{bmatrix}
        SF & | & e'
    \end{bmatrix}
\]
Possible Choice:
\[
    P = \begin{bmatrix}
        I & | & 0
    \end{bmatrix} \ \ \ P' = \begin{bmatrix}
        [e']_\times F & | & e'
    \end{bmatrix}
\]
Canonical Rep:
\[
    P = \begin{bmatrix}
        I & | & 0
    \end{bmatrix} \ \ \ P' = \begin{bmatrix}
        [e']_\times F + e'v^\top & | & \lambda e'
    \end{bmatrix}
\]


\subsection{Struct from imgs}
compute points, triangulate position.

\subsection{Linear Triangulation}
(slide 49-51)\\
multiplicative and additive relationships in $A$
\\
not best for geometric error, good for algebraic error; 
sufficient to show theres a computational solution for computing 3d


\subsection{Ambiguities}
\textbf{Reconstruction uncertainty} - angles/rays close to colinear $\to$ less
specific solution space



\subsection{Refining truct and motion}
Minimize reprojection error
\[
    \min_{\hat P _k, \hat M_i} \sum_{k=1}^{m}\sum_{j=1}^{n}
    D (m_{kj}, \hat P_k \hat M_i)^2
\]

\subsection{Proj. ambiguity + self calibration}
autocalibration (self cal.) - find proj $T$ that upgrades proj. rec. to metric 



\subsection{Orthographic Factorization}
first 2 rows = im1, next 2 = im2, \dots
\\
trackers on points, move camera, etc.
\\
3 first rows in M - 3d structure

\subsection{weak perspective}
M changes by scale, do same thing as orthog but constraints involve scale 



\section{Lecture - Mar 7}
\subsection{Factorization Method Derivation}
2 Steps:
\begin{enumerate}
    \item Show Existence/Mathematical Structure of Prblm
    \item How do we apply to data
\end{enumerate}
(see notebook)
\begin{flalign*}
    \begin{bmatrix}
        U\\V\\W
    \end{bmatrix} &=
    \begin{bmatrix}
        1 & 0& 0&0 \\
        0 & 1 & 0 & 0\\
        0 & 0 & 0 & 1
    \end{bmatrix} 
    \begin{bmatrix}
        \mathbf R & \mathbf t \\
        \mathbf 0 & 1
    \end{bmatrix}
    \begin{bmatrix}
        X \\ Y \\ Z \\ 1
    \end{bmatrix} \\
    \begin{bmatrix}
        u \\ v\\ w
    \end{bmatrix} &=
    \begin{bmatrix}
        r_1^\top & t_1\\
        r_2^\top & t_2 \\
        0^\top & 1
    \end{bmatrix}\begin{bmatrix}
        X \\ Y \\ Z
    \end{bmatrix} =
    \begin{bmatrix}
        r_1^\top \\ r_2^\top
    \end{bmatrix}\begin{bmatrix}
        X \\ Y \\ Z
    \end{bmatrix} + \begin{bmatrix}
        t_1 \\ t_2
    \end{bmatrix}
    % 
    \intertext{Zero mean data $\to$ remove translation; centered coord:}
    \begin{bmatrix}
        \vec u_{ij} \\ \vec v_{ij}
    \end{bmatrix} = 
    \begin{bmatrix}
        \vec u_{ij} - \text{mean $u$} \\ \vec v_{ij} - \text{mean $v$}
    \end{bmatrix}
    \intertext{Const. equation}
\end{flalign*}
Let
\[
    \vec m_{ij} = \begin{bmatrix}
        u_{ij} \\ v_{ij}
    \end{bmatrix} =\begin{bmatrix}
        r_1 ^\top \\ r_2^\top 
    \end{bmatrix} \begin{bmatrix}
        X \\ Y \\ Z
    \end{bmatrix} = [P] [M_j]
\]
\begin{flalign*}
    \intertext{Many pts, $Im_1$}
    \begin{bmatrix}
        u_{11} & u_{12} & \cdots & u_{1n} \\
        v_{11} & v_{12} & \cdots & v_{1n}
    \end{bmatrix} &= P_1 \begin{bmatrix}
        M_1 & M_2 & \cdots & M_n
    \end{bmatrix}
    \intertext{$Im_2$}
    \begin{bmatrix}
        u_{21} & u_{22} & \cdots & u_{2n} \\
        v_{21} & v_{22} & \cdots & v_{2n}
    \end{bmatrix} &= P_2 \begin{bmatrix}
        M_1 & M_2 & \cdots & M_n
    \end{bmatrix}\\
    &\vdots
    \intertext{$Im_m$}
    \begin{bmatrix}
        u_{m1} & u_{m2} & \cdots & u_{mn} \\
        v_{m1} & v_{m2} & \cdots & v_{mn}
    \end{bmatrix} &= P_m \begin{bmatrix}
        M_1 & M_2 & \cdots & M_n
    \end{bmatrix}
\end{flalign*}
\noindent
Put Together
\[
    \overline m = \begin{bmatrix}
        u_{11} & \cdots \\
        v_{11} & \cdots \\
        u_{21} & \cdots \\
        v_{21} & \cdots \\
    \end{bmatrix}^n_{2m} = \begin{bmatrix}
        P_1 \\ P_2 \\ \vline \\ P_m
    \end{bmatrix} \begin{bmatrix}
        M_1 & M_2 & \cdots & M_n
    \end{bmatrix}^n
\]
Showed in rank 3
\\
Perform SVD on $\overline m$ 
\begin{flalign*}
    \overline m &=USV^\top  \\ 
    &= \tilde P \tilde M \text{ where $\tilde P = U$ and $\tilde M = SV^\top$} \\
    &= (PQ) (Q^{-1}M) = \overline P \overline M \text{  ($\tilde P \ne \overline P$, $\tilde M \ne \overline M$; diff coords)}
\end{flalign*}\noindent
We compute 3D structure and cameras up to an affine coordinate transformation 
\\
Rectify Affine to Euclidean:
\begin{flalign*}
    \intertext{Selected}
    [\tilde m] &= Q \underset{\text{(eucl.)}}{\overline M } = Q \begin{bmatrix}
        0 & 0 & \cdots \\
        0 & 2 & \cdots \\
        0 & 0 & \cdots 
    \end{bmatrix}
    \intertext{Solve for $Q$}
    \tilde M &= Q \overline M \\
    ()^\top &= ( \ )^\top \\
    \tilde M ^\top &= \overline M ^\top Q ^\top
    \intertext{Solve LU fact. for $\overline M$}
\end{flalign*}
\noindent
With no 3D ground truth, what do we know $\overline P$ $\gets$ desired
\[
    \overline P = Q \tilde P
\]
rot matrix
\[
    \overline P = \begin{bmatrix}
        \cdots & r_1 & \cdots \\
        \cdots & r_2 & \cdots \\
        & \vdots &
    \end{bmatrix}
\]
Get cholesky decomp.
\begin{flalign*}
    0 = r_1^\top r_2 &= P_1 Q^\top Q P_2 \\
    &= P_1 C P_2\\
    s^2 = r_1^\top r_1 &= C \\
    s^2 = r_2^\top r_2 &= C \\
    & \downarrow
    \\
    & \cdots
\end{flalign*}
% \subsubsection*{Step 1: Show Existence/Mathematical Structure of Prblm}
% \subsubsection*{Step 2: How do we apply to data}







\end{document}

